Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-18 paper/paper.tex
added citations Tested everything on quipu.

\section{Introduction}
\fixme{Huang and those who reference him use the terms 'numerical
  deterministic reaction-diffusion model' and 'stochastic version of
  this model' for the stochastic.} It is vital that during the process
of bacterial cell division a cell avoid minicelling, or splitting into
daughter cells with lopsided volumes.  Instrumental to this process in
\emph{Escherichia coli} is a long FtsZ polymer chain that develops on
the cell wall in the center region of the cell, helping dictate the
plane of
division~\cite{adams2009bacterial,lutkenhaus2007assembly}. Previous
experimental studies have shown that the MinC protein, known to
inhibit the FtZ polymer\cite{shen2010examination}, exhibits regular
pole to pole oscillatory behavior between both ends of the wild-type
pill-shaped cell.  It thus has a higher time averaged concentration in
the cell poles than in the center region, which aides in prohibiting
the FtZ from developing in the wrong region.  The MinC is recruited to
these poles by MinD, which ineracts with another protein, MinE, that
is neccesary for MinD oscillation.~\cite{hu1999topological,
  fu2001mine, shapiro2009and, yu1999ftsz, raskin1999rapid}.

Previous studies have shown that the minD protein system is capable of
exhibiting oscillations in round shapes \fixme{citation needed}as well
as in connected three pronged tube shapes~\cite{varma2008min}, in
which the oscillations seem to seek out the extreme poles in the
cell~\cite{corbin2002exploring} \cite{juarez2010changes}. However,
Mannik \emph{et al.} have also shown that there are limitations to
this capability~\cite{mannik2010bacteria}
\cite{mannik2009bacterial}. They have experimentally forced E. Coli
cells into flattened irregular shapes of $0.25\mu m$
thickness. Oscillations in this type of cell shape lose their
regularity, both temporally and spatially. This allows for an
opportunity to test previously studied MinD simulation models against
more extreme experimental cases than have been seen thus
far. \fixme{Perhaps this sentence should come later after we've
  introduced the simulation research that's been done.}

>>>>

\section{Introduction}
\fixme{Huang and those who reference him use the terms 'numerical
  deterministic reaction-diffusion model' and 'stochastic version of
  this model' for the stochastic.} It is vital that during the process
of bacterial cell division a cell avoid minicelling, or splitting into
daughter cells with lopsided volumes.  Instrumental to this process in
\emph{Escherichia coli} is a long FtsZ polymer chain that develops on
the cell wall in the center region of the cell, helping dictate the
plane of
division~\cite{adams2009bacterial,lutkenhaus2007assembly}. Previous
experimental studies have shown that the MinC protein, known to
inhibit the FtZ polymer\cite{shen2010examination}, exhibits regular
pole to pole oscillatory behavior between both ends of the wild-type
pill-shaped cell.  It thus has a higher time averaged concentration in
the cell poles than in the center region, which aides in prohibiting
the FtZ from developing in the wrong region.  The MinC is recruited to
these poles by MinD, which ineracts with another protein, MinE, that
is neccesary for MinD
oscillation.~\cite{hu1999topological,fu2001mine,shapiro2009and,yu1999ftsz,raskin1999rapid,meacci2005min,raskin1999minde}.

Previous studies have shown that the minD protein system is capable of
exhibiting oscillations in round shapes \cite{fange2006noise}as well
as in connected three pronged tube shapes~\cite{varma2008min}, in
which the oscillations seem to seek out the extreme poles in the
cell~\cite{corbin2002exploring} \cite{juarez2010changes}. However,
Mannik \emph{et al.} have also shown that there are limitations to
this capability~\cite{mannik2010bacteria}
\cite{mannik2009bacterial}. They have experimentally forced E. Coli
cells into flattened irregular shapes of $0.25\mu m$
thickness. Oscillations in this type of cell shape lose their
regularity, both temporally and spatially. This allows for an
opportunity to test previously studied MinD simulation models against
more extreme experimental cases than have been seen thus
far. \fixme{Perhaps this sentence should come later after we've
  introduced the simulation research that's been done.}
